---
title: "KzSynITScom_metadata_prep"
author: "Charles T. Bond"
date: "10/25/2023"
output: html_document
---
### Setup
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/chunk/AIMS_Konza_synoptic_ITS_community/metaprep")
knitr::opts_knit$set(root.dir = "/Users/chunk/AIMS_Konza_synoptic_ITS_community/metaprep")
```

```{r, library}
library(tidyverse)
library(RColorBrewer)
#library(ggpubr)
```


# Metadata preparation
We are pulling from multiple published sources for our metadata, which are published in part on the AIMS project hydroshare page (https://www.hydroshare.org/group/247) listed as follows:

### Site physical characteristics
Ramos, R., A. Burgin, S. Zipper (2023). Konza_Synoptic_Physical_Characteristics, HydroShare, http://www.hydroshare.org/resource/95c18f7643534def99d40b7b07971bee


### Chlorophyll-a data
Bond, C. T., K. A. Kuehn, E. Stanley (2023). Konza_Synoptic_Chorophyll, HydroShare, http://www.hydroshare.org/resource/01135480279340cd8e457a22e7b9208b

### Dissolved gasses
link tbd

### Extracellular Enzymes
link tbd

### StreamDAG


```{r}
## Field datasheet data on Leaf data, and long-term STIC data, are in the following table 'meta_base'
setwd("/Users/chunk/AIMS_Konza_synoptic_ITS_community/metaprep")
meta_base <- read.csv("meta_base_kzsynITS.csv")
#View(meta_base)

### AIMS Konza synoptic Physical Characteristics table was published seperately
Konza_Synoptic_Physical_Characteristics <- read.csv("/Users/chunk/AIMS_Konza_synoptic_ITS_community/metaprep/Konza_Synoptic_Physical_Characteristics.csv")

#merge by siteid
meta_tab<- merge(meta_base,Konza_Synoptic_Physical_Characteristics, by="siteid", all=TRUE)
```



placeholder stream DAG example
```{}
# newest streamDAG
#library(devtools)
#install_github("moondog1969/streamDAG")
library(streamDAG)

data(AIMS.node.coords)
td_graph <- streamDAGs("td_full")
td_coords <- AIMS.node.coords[AIMS.node.coords$site == "TD",]
with(td_coords, spatial.plot(td_graph, long, lat, Object.ID,  
main = "Talladega"))



set.seed(20)
fake.td.lengths <- rnorm(length(E(td_graph)), 30, 8)

td_graphw <- td_graph
E(td_graphw)$weight <- fake.td.lengths
sia <- size.intact.to.arc(td_graphw)

td_graphw2 <- td_graph
E(td_graphw2)$weight <- sia
ac <- local.summary(td_graphw2, "alpha.cent") 



```

```{r}
library(streamDAG)
data(AIMS.node.coords)
kon_full <- streamDAGs("konza_full")

data(kon_coords)
spatial.plot(kon_full, kon_coords[,3], kon_coords[,2], names = kon_coords[,1])
A.mult(kon_full, power = 6, text.summary = TRUE)

data("kon_lengths")


# now, we want to set nodes at dry sites as absent in order account for disconnections in the network in our local network connectivity metrics, and then we want to weigh alpha centrality ('node importance') by flowing upstream network length...

meta_tab$wetdrybin <- ifelse(meta_tab$flow_state == "dry", 0, 1)
meta_tab$wetdrybin


##there is probably a shorter way to get the wet/dry binary vector for the 46 sites, but this does the job. 
local <- as.data.frame(local.summary(kon_full))
t.local<- as.data.frame(t(local)) %>%
  rownames_to_column(var="siteid")
t.local$siteid <- substr(t.local$siteid, 1, nchar(t.local$siteid) - 2)
t.local$nrow<- 1:nrow(t.local)
temptab<- merge(meta_tab,t.local, by="siteid") ### all = FALSE give just the 46 sites, TRUE give all 50.
temptab<- temptab[order(temptab$nrow), ]
#meta_tab[meta_tab$wetdrybin==0,]
#tempsite<- temptab[temptab$wetdrybin==0,]
#tempsite$siteid
#temptab[temptab$flow_state=='dry',]

drybin<- temptab$wetdrybin
temptab$wetdrybin
drybin
### weird glitch, rechecked and tried rerun multiple times but the computer keeps switching 04M13 back to wet???? Manually setting as dry (it is the first site in order on kon_full object)
npbin<- temptab$notpoolbin
temptab$notpoolbin
npbin


kon_graphw <- kon_full


### now, because the node for 20M03 is missing (which happened to be dry), the arc for 20M04_1->20M02_1 was not deleted, despite 20M02 being a disconnected pool at the time we sampled. I will manually remove te arc for 20M04_1->20M02_1.

#rm.arcs.bin<- ifelse(E(kon_graphw) == "20M04_1->20M02_1", 0, 1) ### did
rm.arcs.bin <- c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1)

kon_graphw<- delete.arcs.pa(kon_graphw,rm.arcs.bin)
kon_graphw

kon_graphw<- delete.nodes.pa(kon_graphw,drybin)
kon_graphw


spatial.plot(kon_graphw, kon_coords[,3], kon_coords[,2], plot.dry = TRUE, names = kon_coords[,1])
wetlength<- kon_lengths[E(kon_graphw),]

#kon_graphw<- delete.edges(kon_graphw,drybin)
E(kon_graphw)$weight <- wetlength$Lengths
sia <- size.intact.to.arc(kon_graphw)
sin<- size.intact.to.node(kon_graphw, node="all")



kon_graphw2 <- kon_graphw
E(kon_graphw2)$weight <- sia
E(kon_graphw2)$weight
ac <- local.summary(kon_graphw2, "alpha.cent") 
#ac <- local.summary(kon_graphw2, "alpha.cent") 

##Local Measures for 46/50 of the transects included in this stude
local <- as.data.frame(local.summary(kon_graphw))
t.local<- as.data.frame(t(local)) %>%
  rownames_to_column(var="siteid")
t.local$siteid <- substr(t.local$siteid, 1, nchar(t.local$siteid) - 2)

t.local$alpha.cent.wt<- log(1+ac)
t.local$flowing.upstream.length.m<- sin
head(t.local)
plot(t.local$flowing.upstream.length.m,t.local$alpha.cent)
plot(log(1+t.local$flowing.upstream.length.m),log(1+t.local$alpha.cent.wt))
```

Merge stream dag local measures to meta_tab...
```{r}
meta_tab<- merge(meta_tab,t.local, by="siteid", all=TRUE)

```

Great, now add chlorophyll, EEA, and qPCR results 
```{r}
## load chlorophyll data
Chla<- read.csv("/Users/chunk/AIMS_Konza_synoptic_ITS_community/metaprep/CHLA_GP_approach3_20210605_20210607.V1.0.csv")
#merge by siteid
meta_tab<- merge(meta_tab,Chla, by="siteid", all=TRUE)


### load enzyme data
EEA<- read.csv("/Users/chunk/AIMS_Konza_synoptic_ITS_community/metaprep/KzSyn_EEA_overall_7_7_23.csv")
#merge by siteid
meta_tab<- merge(meta_tab,EEA, by="siteid", all=TRUE)

### load enzyme data
qpcr<- read.csv("~/AIMS_Konza_synoptic_ITS_community/metaprep/QPCR_GP.csv")
#merge by siteid
meta_tab<- merge(meta_tab,qpcr, by="siteid", all=TRUE)

```

Done with the meta_data?
```{r}

write.csv(meta_tab, "/Users/chunk/AIMS_Konza_synoptic_ITS_community/metatab111123.csv")

```

Now, since each site can have multiple as many as 4 substrates with sequencing results, we need to make the table work with our phyloseq object downstream. 
```{r}

substrates<- read.csv("/Users/chunk/AIMS_Konza_synoptic_ITS_community/kz.syn.ITS.substrates.csv")
#merge by siteid


meta_tab<- merge(substrates,meta_tab, by="siteid", all=TRUE)
rownames(meta_tab)<-meta_tab$Sample

write.csv(meta_tab, "/Users/chunk/AIMS_Konza_synoptic_ITS_community/submetatab111123.csv")

```

##Konza map
Now, downloaded stream shapefile from here: http://www.konza.ksu.edu/map/
```{r}

# for loading our data
#library(jsonlite)
library(rgdal)
library(sf)
# for plotting
#library(extrafont)
#library(ggplot2)
#library(ggspatial)
#library(patchwork)
#library(scico)
#library(vapoRwave)
# for data wrangling
#library(dplyr)
#konza_strs<- read_sf("GIS210.shp")

kzmap<- st_read("GIS210.shp")

kzmap_ll <- st_transform(kzmap, "+proj=longlat +ellps=WGS84 +datum=WGS84")

#head(st_coordinates(kzmap_ll))
# Check the extent of the transformed data
print(st_bbox(kzmap_ll))

# Check the CRS of the transformed data
print(st_crs(kzmap_ll))

strmp1 <- ggplot(data = kzmap_ll) +
  geom_sf() +
  coord_sf(ylim=c(39.0700, 39.095000), xlim = c(-96.60000, -96.56500))
strmp1

## percent wet color scale, same as for dbRDA plots
pal<- c("#046C9A","#3A9AB2","#85B7B9","#ADC397","#DCCB4E","#E5A208","#ED6E04","#F11B00") # too red
common_limits <- c(0, 1) 
sticwet<- ggplot(data=meta_tab, aes(x=long, y=lat, colour=prc_wet))+
  geom_point(size=3.5)+
  scale_colour_gradientn(colors = rev(pal), labels = scales::percent_format(), limits = common_limits)
sticwet

mapmap <- ggplot() +
  geom_sf(data = kzmap_ll) +
  coord_sf(ylim=c(39.07, 39.095), xlim = c(-96.595, -96.565))+
  geom_point(data=meta_tab, aes(x=long, y=lat, colour=prc_wet, shape=flow_state), size=3.3)+
  scale_colour_gradientn(colors = rev(pal), labels = scales::percent_format(), limits = common_limits)+
  labs(legend="right" ,shape="Flow state:", color="Percent wet:", x="Longitude", y="Latitude"
      )+ 
 theme(axis.text.x = element_text(size = 8), axis.text.y = element_text(size = 8))
mapmap



```


